---
title: "BPC Prostate - Genomic descriptives"
author: "Alex Paynter"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F,
                      warning = F, message = F,
                      fig.width = 7, fig.height = 5)
k_dpi <- 150
pal_pt_hc <- c("#004488","#ddaa33", "#bb5566")
```

<style type="text/css">
.main-container {
  max-width: 1000px !important;
  margin: auto;
}
</style>

```{r}
library(purrr)
library(here)
library(fs)

purrr::walk(.x = fs::dir_ls(here('R')), .f = source) # also loads lots of packages.
```



```{r, set_gtsummary_theme}
theme_gtsummary_compact()
theme_gtsummary_language("en", big.mark = "") # for now.


```


```{r, warning = T}
read_wrap_clin <- function(p) {
  read_rds(file = here("data", 'clin', p))
}

dft_pt <- read_wrap_clin("dft_pt.rds")
dft_ca_ind <- read_wrap_clin("dft_ca_ind.rds")
dft_cpt <- read_wrap_clin("dft_cpt.rds")

read_wrap_geno <- function(p) {
  read_rds(file = here("data", 'genomic', p))
}

dft_gp_all <- read_wrap_geno('gene_panel_all.rds')
dft_gp_sum <- read_wrap_geno('gene_panel_sum.rds')
dft_gp_by_gene <- read_wrap_geno('gene_panel_by_gene.rds')
dft_onco_impact <- read_wrap_geno('oncokb_impact.rds')

dft_alt <- read_wrap_geno('alterations.rds')

dft_gt_any_alt <- read_wrap_geno(
  'gene_test_any_alt.rds'
)


```


## Content


```{r}
# Table explaining the sample type coding we'll use:
dfp_sample_type_expl <- dft_cpt %>% 
  group_by(sample_type) %>%
  summarize(
    num_sample = n(),
    # Each group should have one unique sample_type_simple_f - but if not this will fail loudly in the table. 
    coded_as = paste(unique(sample_type_simple_f)),
    .groups = "drop"
  ) %>%
  arrange(desc(num_sample))

# pp = per person.
dft_ngs_pp <- dft_cpt %>%
  group_by(record_id) %>%
  summarize(
    All = n(),
    Primary = sum(sample_type_simple_f %in% "Primary tumor"),
    Metastatic = sum(sample_type_simple_f %in% "Metastatic")
  ) 

dft_ngs_pp_long <- dft_ngs_pp %>%
  pivot_longer(
    cols = -record_id,
    names_to = "category",
    values_to = "count"
  ) %>%
  mutate(
    category = forcats::fct_inorder(category)
  )

gg_ngs_pp <- ggplot(
  data = dft_ngs_pp_long,
  aes(x = count, fill = category)
) + 
  geom_histogram(binwidth = 1, center = 0) +
  theme_classic() + 
  labs(x = "Number of tests",
       y = "Participants") + 
  scale_y_continuous(expand = expansion(add = 0, mult = c(0,0.05)),
                     n.breaks= 8) +
  scale_x_continuous(
    breaks = 0:10,
    expand = expansion(
      add = c(0,0), 
      mult = c(0,0.01)
    )
  ) + 
  facet_wrap(vars(category)) +
  theme(
    strip.text = element_text(hjust = 0),
    legend.position = "none"
  ) + 
  scale_fill_manual(
    values = pal_pt_hc
  )

n_prim_met_pair <- dft_ngs_pp %>% 
  filter(Primary >= 1 & Metastatic >= 1) %>%
  nrow(.)

n_gte_2_prim <- dft_ngs_pp %>% 
  filter(Primary >= 2) %>%
  nrow(.)

n_gte_2_met <- dft_ngs_pp %>% 
  filter(Metastatic >= 2) %>%
  nrow(.)
                            
```


### R2.1.1 NGS tests per person

This table uses the "Specimen Sample Type" variable from the Cancer Panel Test dataset.  We simplified this slightly (`coded_as`) to avoid small categories of tumors:

```{r, include = T}
dfp_sample_type_expl %>%
  huxtable(.) %>%
  huxtable::set_align(value = "left") %>%
  theme_compact(.)
```

The following plot shows histograms of the number of tests per participant.  This is repeated for all tests (Primary, Metastatic, Other) and separately for Primary and Metastatic tumors.

```{r, include = T, fig.height = 3}
gg_ngs_pp
```

Observations:

- The overwhelming majority of participants have exactly one NGS test.
- About 400 people have no primary tumor tests at all.
- `r n_prim_met_pair` people have at least one primary and at least one metastatic tumor.
- Only `r n_gte_2_prim` participants have more than one primary tumor test.
- There are `r n_gte_2_met` with more than one metastatic test, overwhelmingly from MSK.










```{r}
gg_gene_sample_hist <- ggplot(
  dft_gp_by_gene,
  aes(x = prop_samp_tested)
) +
  geom_histogram(binwidth = 0.1, center = 0.05, fill = '#6699cc') +
  theme_classic() + 
  labs(x = "Proportion of samples gene tested in",
       y = "Number of genes") + 
  scale_y_continuous(expand = expansion(add = 0, mult = c(0,0.05)),
                     n.breaks= 8) +
  scale_x_continuous(
    labels = scales::percent,
    expand = expansion(add = c(0,0), mult = c(0,0.01)),
    breaks = seq(0,1, by = 0.1)) + 
  theme(
    # jump up the right margin to get the '100%' in 
    plot.margin = margin(unit(c(5.5, 10, 5.5, 5.5), "points"))
  )

# Quick computation for the observations below:
n_genes_total_coverage <- dft_gp_by_gene %>% filter(prop_samp_tested >= 0.999) %>% nrow()
n_genes_90_cov <- dft_gp_by_gene %>% filter(prop_samp_tested >= 0.899) %>% nrow()
n_genes_70_cov <- dft_gp_by_gene %>% filter(prop_samp_tested >= 0.699) %>% nrow()
dft_gp_by_gene %>% filter(prop_samp_tested < 0.5) %>% nrow()

```

### R2.1.2 Gene NGS coverage over samples

There are `r length(unique(dft_gp_by_gene$hugo))` genes covered by at least one panel in this dataset.  This is a histogram of genes, showing the number covering each decile of the proportion of samples (see example interpretation below).

```{r, include = T, fig.height = 4}
gg_gene_sample_hist
```

**Example interpretation:** The bar between 90% and 100% shows about 200 genes.  This means that approximately 200 genes were covered in at least 90% of panels.  Likewise, there were nearly 300 genes which were covered in under 10% of samples.

Observations:

- There are only `r n_genes_total_coverage` genes covered by every panel.  For our clinico-genomic analyses analyses it might be better to include genes covered by **most** panels to gather a larger set.  Two natural cut points based on the graph above may be:
  - Genes covered by at least 90% of panels: `r n_genes_90_cov`.
  - Genes covered by at least 70% of panels: `r n_genes_70_cov`.
  - Keep in mind that some of these will have too few positives to be used, further reducing the numbers of genes included.  We'll describe this in subsequent sections.
- The majority of genes are covered by under 50% of samples.  These will almost certainly need to be excluded for clinico-genomic analyses due to the degree of missingness.








```{r}
# Add the number of patients and samples in parens:
dft_gp_all %<>%
  arrange(cpt_seq_assay_id) %>%
  left_join(
    ., 
    select(dft_gp_sum, cpt_seq_assay_id, n_pts, n_samples, n_genes),
    by = "cpt_seq_assay_id"
  ) %>%
  mutate(
    assay_lab = glue("{cpt_seq_assay_id} [{n_genes}] ({n_pts}, {n_samples})"),
    assay_lab = forcats::fct_inorder(assay_lab)
  ) 

# Get the list of genes covered in >50% of samples and which have at least 1 alteration in the data:

vec_gene_gte1_alt <- readr::read_rds(
  here('data', 'genomic', 'gene_gte1_alt.rds')
)
dft_gene_prop_samp_test <- readr::read_rds(
  here('data', 'genomic', 'gene_prop_samp_test.rds')
)
vec_test_prop_gte50 <- dft_gene_prop_samp_test %>%
  filter(prop >= 0.5) %>%
  pull(hugo)
dft_gp_all_limited <- dft_gp_all %>%
  filter(hugo %in% vec_gene_gte1_alt) %>%
  filter(hugo %in% vec_test_prop_gte50)


gg_panel_coverage <- plot_assay_heatmap(
  dat = dft_gp_all,
  plot_title = "Gene panel coverage"
)

gg_panel_coverage_limited <- plot_assay_heatmap(
  dat = dft_gp_all_limited,
  plot_title = "Gene panel coverage (limited)"
)
  
  
  
```


### R2.1.3a Panel coverage

Heatmap of gene panel coverage.

```{r, include = T, fig.width = 10, fig.height = 5}
gg_panel_coverage
# Technically this works but it takes too long:
# plotly::ggplotly(gg_panel_coverage)
```

Observations:

- Panels vary in size considerably (under 50 to over 500). 
- Currently this figure is order by number of patients on the Y axis.  It might be more informative to cluster the axis by similarity.  Some obvious groups of panels jump out to me:
  - {DFCI-ONCOPANEL-3, DFCI-ONCOPANEL-3.1, UNH-555-*}
  - {MSK_IMPACT468, DFCI-ONCOPANEL-2, VICC-01-T7}
  - The two UHN panels with few genes covered. 







```{r}
n_pt <- dft_ca_ind %>% nrow
n_samp <- dft_cpt %>% nrow

gene_panel_table_help <- function(dat) {
  dat %>% 
    mutate(
      Patients = glue("{n_pts} ({round(n_pts/n_pt*100, 1)}%)"),
      Samples = glue("{n_samples} ({round(n_samples/n_samp*100, 1)}%)"),
      Genes = n_genes
    ) %>%
    select(
      Panel = cpt_seq_assay_id,
      Patients,
      Samples,
      Genes
    ) %>%
    distinct
}

dfp_gp <- gene_panel_table_help(dft_gp_all)
dfp_gp_limited <- gene_panel_table_help(dft_gp_all_limited)

gp_table_hux <- function(dat) {
  dat %>%
    huxtable(.) %>%
    huxtable::set_align(value = "left") %>%
    theme_compact(.)
}
  
```

**Table Version**  The following table shows similar information to the above figure: the number of patients, samples and genes covered by each panel.  

```{r, include = T}
gp_table_hux(dfp_gp)
```

*Note:* The patient percentages do not add to 100 (because each person can be tested with multiple panels).



```{r}
genes_panel <- dft_gp_all %>% 
  pull(hugo) %>% 
  unique %>%
  length


genes_panel_limited <- dft_gp_all_limited %>% 
  pull(hugo) %>% 
  unique %>%
  length
```


### *R2.1.3b Panel coverage, limited

Same as R2.1.2a except we limit to genes with at least one alteration in the dataset, and genes that are covered in the panels used in at least 50% of samples.  This selection brings us down to `r genes_panel_limited` genes, from the original `r genes_panel`.

```{r, show_gene_figure_limited, include = T}
gg_panel_coverage_limited
```

The table version of the limited set above would be the same except for the `Genes` column.  Skipping for now.









```{r}
dft_onco_impact %<>% 
  mutate(n_div = if_else(
    oncogenic %in% c("Oncogenic", "Likely Oncogenic"),
    n,
    -n)
  ) %>%
  arrange(type, oncogenic)

# hcl_palettes("diverging", n = 5, plot = T)
pal_onco <- diverging_hcl(
  palette = "Tropic", n = 5
)




gg_onco_impact <- ggplot(
  dft_onco_impact,
  aes(x = n_div, fill = oncogenic, y = type)
) + 
  geom_col(
    position = "stack", orientation = 'y'
  ) +
  scale_fill_manual(
    values = pal_onco,
    name = NULL
  ) +
  guides(
    fill = guide_legend(reverse = T)
  ) + 
  labs(
    x = "Alterations (bulk count)",
    title = "Oncogenic annotation with OncoKB"
  ) + 
  theme_classic() + 
  theme(
    legend.position = "bottom",
    plot.title.position = "plot"
  )
```




### R2.2.1 Oncogenic annotation impact

We discussed annotating and potentially filtering the data using the "oncogenic" feature from oncoKB (https://www.oncokb.org/).  Looking at the alteration level (multiple alterations per sample are possible), the following plot shows the impact this would have on the raw data.

```{r, include = T, fig.height = 4}
gg_onco_impact
```

Observations:

- If we filtered our data to include only likely oncogenic or oncogenic alterations, we lose most of the mutations, and a smaller proportion or the fusions and CNAs.







### R2.2.2 Onco. impact by pathways

```{r}

```









```{r}
dft_inst_freq <- dft_gt_any_alt %>%
  select(
    sample_id,
    hugo,
    tested,
    any_alt,
    any_alt_onco
  ) %>%
  left_join(
    .,
    select(
      dft_cpt, 
      sample_id = cpt_genie_sample_id,
      institution
    ),
    by = "sample_id",
    relationship = 'many-to-one'
  )

dft_inst_freq %<>%
  group_by(institution, hugo) %>%
  summarize(
    n_tested = sum(tested, na.rm = T),
    n_alt = sum(any_alt, na.rm = T),
    n_alt_onco = sum(any_alt_onco, na.rm = T),
    
    prop_alt = n_alt/n_tested,
    prop_alt_onco = n_alt_onco/n_tested,
    .groups = "drop"
  ) 

dft_inst_freq_all <- dft_inst_freq %>%
  group_by(hugo) %>%
  summarize(
    across(
      .cols = c(n_alt, n_alt_onco, n_tested),
      .fns = sum
    ),
    .groups = 'drop'
  ) %>%
  mutate(
    prop_alt = n_alt/n_tested,
    prop_alt_onco = n_alt_onco/n_tested
  ) %>%
  mutate(institution = "All")
    

# find the genes that are most freqently altered, limited
# to those with over 600
vec_alt_most_common <- dft_inst_freq_all %>%
  filter(n_tested > 0.5*max(n_tested)) %>%
  arrange(desc(prop_alt)) %>%
  slice(1:20) %>%
  pull(hugo)

dfp_inst_freq <- bind_rows(
  dft_inst_freq_all,
  dft_inst_freq
) %>%
  filter(hugo %in% vec_alt_most_common) %>%
  mutate(
      institution = forcats::fct_inorder(institution),
      hugo = factor(hugo, levels = vec_alt_most_common)
  )



gg_inst_alt <- plot_alt_by_inst(
  dat = dfp_inst_freq,
  point_size = 2
)

gg_inst_alt_onco <- plot_alt_by_inst(
  dat = dfp_inst_freq,
  prop_var = "prop_alt_onco",
  plot_subtitle = "Only oncogenic alterations",
  point_size = 2
)

```
  
### *R2.3.1a Alteration frequency by institution

The following plot show the percentage of tested samples with an alteration of any kind.  The black square shows the whole cohort number, and colored dots show the numbers at each institution.  

This plot is limited to genes tested in more than 50% of samples, and only the top 20 most frequently altered genes are shown.

```{r, include = T}
gg_inst_alt
```

**Observations:**

- Some genes are missing dots for institutions because the gene was never tested there.  An example is MGA at VICC and UHN - not included in any panel in this dataset.
- Most of the big outliers are at the small institutions.
- Even at DFCI and MSK, proportions of altered samples can be quite different depending on the gene.  An example that remains in the next plot is MYC, where DFCI has 2% of tested samples altered and MSK has 9%.

### *R2.3.1b Alteration frequency by institution (oncogenic)

Same as the above plot, except we examine include only alterations which were annotated as Likely Oncogenic or Oncogenic by OncoKB.



```{r, include = T}
gg_inst_alt_onco
```









## Upcoming

- Further charcterization of alterations after we decide on filtering:
  - Consequence (missense, frameshift, synonymous, etc).
  - Level of evidence (for likely oncogenic or oncogenic mutations).
- Lookup table for genes.
- Co-occurence
- Pathways





## Problems/Concerns

### Structural variants

**Question:** At one of our early meetings, someone asked about structural variants and whether they would be covered by our assessments. 

**Answer:** Yes, with some limits.  In main GENIE v12, which forms the foundation for this BPC cohort, fusion files were still in use and had not been replaced by SV files.  We will use this file for analysis, and while we know the class of structural variants is broader than fusions, my understanding is that in practice these pipelines/files cover similar ground.

There is also the question of whether a particular panel covers structural variants - some do and some do not in this dataset.  The vast majority of our samples are tested with panels that can assess structural variants (for example, all the MSK IMPACT and DFCI ONCOPANELs look at SVs).  Some of the smaller panels (UHN_48-V1) do not cover SVs.  

A similar issue is present for CNAs - some panels cover them and some don't.  Additionally, reading the data guide, some only cover gene level CNAs and not intragenic CNAs. 







