---
title: "BPC Prostate - Abi/Enza/Doce"
author: "Alex Paynter"
date: "`r Sys.Date()`"
output:
  quarto::quarto_html:
editor_options:
  quarto:
    chunk_output_type: console
format:
  html:
    embed-resources: true
    toc: true
    theme: sandstone 
execute:
  echo: false
  include: false
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---

```{r}
library(purrr); library(here); library(fs)
purrr::walk(.x = fs::dir_ls(here('R')), .f = source)
```


```{r}
js_sun_simple <- readr::read_rds(
  file = here('data', "sunburst_plots", "abi_enza_doce_simple.rds"),
)
```


## Introduction

This is a version of a plot from the Prostate cancer report with one major change: We exclude regimens that contain more than one drug in the set under consideration.  For example, if a regimen has both Abiraterone and Docetaxel, we exclude that regimen.  This is for simplicity of presentation.

## Plot

This plot studies the set of drugs {Abiraterone Acetate, Docetaxel, Enzalutamide}, and the question is "what order do people tend to take them in?"  The participants included are those who took one of these three drugs at any point in the study.  Repeat regimens, for example Enzalutamide followed by Enzalutamide, are ignored.

```{r, include = T}
js_sun_simple
```



