---
title: "DoD PCRP notes"
subtitle: "Not a part of the manuscript work - for Sage/AACR proposal"
author: "Alex Paynter"
date: "`r Sys.Date()`"
editor_options:
  quarto:
    chunk_output_type: console
format:
  html:
    embed-resources: true
    toc: true
    theme: sandstone 
execute:
  echo: true
  include: true
  warning: false
  message: false
  fig.width: 7
  fig.height: 5
---

```{r}
#| include: false
#| echo: false
library(purrr); library(here); library(fs)
purrr::walk(.x = fs::dir_ls(here('R')), .f = source) # also loads lots of packages.
```


```{r}
#| include: false
#| echo: false
read_wrap_clin <- function(p) {
  read_rds(file = here("data", 'clin', p))
}

dft_pt <- read_wrap_clin("dft_pt.rds")
dft_ca_ind <- read_wrap_clin("dft_ca_ind.rds")
# we're taking the augmented version, which has TMB columns added.  The existing info is the same.
dft_cpt <- read_wrap_clin("dft_cpt_aug.rds")
```

## Introduction

This report is a place to collate numbers for a grant submission.

## A litany of numbers

**Number in the cohort:** Specifically we pull index cases:

```{r}
nrow(dft_ca_ind)
```

**Number and percent metatstatic in the cohort:**

```{r}
n_met <- nrow(get_dmet_time(dft_ca_ind))
n_met
n_met / nrow(dft_ca_ind)
```

A bit more context on what we mean by metastatic people:

```{r}
get_dmet_time(dft_ca_ind, annotate_type = T) %>%
  count(.met_type)
```

All cases in the BPC prostate cohort have at least one next-generation sequencing test. Here's what that means for mutations, CNAs and fusions:


```{r}
#| echo: false
dft_cpt <- read_wrap_clin("dft_cpt_aug.rds")

read_wrap_geno <- function(p) {
  read_rds(file = here("data", 'genomic', p))
}

dft_gp_all <- read_wrap_geno('gene_panel_all.rds')

dft_gp_all %>% 
  group_by(cpt_seq_assay_id) %>%
  summarize(
    any_mut = any(tested),
    any_cna = any(tested_cna),
    any_fusion = any(tested_fusion),
    .groups = "drop"
  ) %>%
  left_join(
    select(dft_cpt, record_id, cpt_seq_assay_id),
    .,
    by = "cpt_seq_assay_id"
  ) %>%
  group_by(record_id) %>%
  summarize(
    any_mut = any(any_mut),
    any_cna = any(any_cna),
    any_fus = any(any_fusion),
    .groups = "drop"
  ) %>%
  summarize(
    people_with_gte_one_sample_testing_mut = n_pct_str(sum(any_mut), nrow(dft_ca_ind)),
    people_with_gte_one_sample_testing_cna  = n_pct_str(sum(any_cna), nrow(dft_ca_ind)),
    people_with_gte_one_sample_testing_sv = n_pct_str(sum(any_fus), nrow(dft_ca_ind))
  ) %>% 
  glimpse


    
```

